Complete Real-Time Liver Model including Glisson Capsule, Vascular Network and Parenchyma

### 2013-02-28 Doc/Miccai2013/miccai2013.tex
ADD and CHANGE: some additions and modifications in Model -- Parenchyma and Vascularization
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@528 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\def\TODO{{\color{red}{TODO}}}

%%% Comments
\usepackage{color}
\newcommand{\TG}[1]{{\color{blue}\textbf{TG: #1}}}
\newcommand{\CD}[1]{{\color{green}\textbf{CD: #1}}}
\newcommand{\IP}[1]{{\color{cyan}\textbf{IP: #1}}}
%%%

@@

The paper is organized as follows: \TODO.

@@

In this section we will describe the construction of our composite model.
It's three main parts are: tetrahedral model for parenchyma, membrane model
for capsule coupled to the parenchyma and beam model of vascular system as
proposed by Peterl\'{i}k et al. \cite{Peterlik2012}.

\subsection{Parenchyma} %{{{

It is known that the parenchyma exhibits viscoelastic behaviour \TG{Cite
something}, we however employ simpler elastic model as we are not so much
interested in time-dependent behaviour but rather in static equilibrium
under certain conditions.
We also rely on the vascularized model of the parenchyma proposed by
Peterl\'{i}k et al. \cite{Peterlik2012}.

The parenchyma is modeled using corotational finite elements. The corotational
formulation allows us to deal with geometrical non-linearities but still
rely on linear stress-strain relationship. Thus rotations and large
deformations are possible. In the corotational formulation the deformation
of each element is expressed in the local frame of reference. There are
several ways of computing the corotational frame for elements, we rely on
the geometrical method proposed in \cite{Nesme2005}.
% NOTE: This description of corotational method is very simplified and could be extended.
The stiffness matrix $\Mat{K}$ can be expressed in terms of the deformation
$u$ and the force as $\Vec{f} = \Mat{K}(\Vec{u}) \Vec{u}$. Where the
components of $\Mat{K}$ depend on the orientation of each element in the
current simulation step.

The vascularization method is based on linear beams
with local frames of reference (in a many aspects similar to the
corotational formulation). As such the model also handles geometric
non-linearities in the deformation. The proposed model assumes no relative
motion of vessels to the parenchyma. Segmented vascular network is
discretized into the set of linear elements that are then projected into
the tetrahedral mesh of the parenchyma. During the simulation forces and
torques resulting from the deformations of vessels by parenchyma are
computed and propagated back onto the tetrahedral mesh.

>>>>

%%% Comments
\usepackage{color}
\newcommand{\TG}[1]{{\color{blue}\textbf{TG: #1}}}
\newcommand{\CD}[1]{{\color{green}\textbf{CD: #1}}}
\newcommand{\IP}[1]{{\color{cyan}\textbf{IP: #1}}}
\newcommand{\TODO}[1]{{\color{red}\textbf{TODO: #1}}}
%%%

@@

The paper is organized as follows: \TODO{paper organization}.

@@

In this section we describe the construction of our composite model
being composed of three main components: tetrahedral FE model of the 
parenchyma, Tymoshenko-beam model for the vascular structures and finally 
triangular membrane elements used for the capsule.

\subsection{Parenchyma and Vascularization} %{{{

It is known that the parenchyma exhibits non-linear viscoelastic behaviour \cite{marchesseauMJED}.
However, as we are mainly interested in the static equilibrium, we do not model the time-dependent
phenomena related to viscosity.

%However, we employ simpler corotational model as we are not so much
%interested in time-dependent behaviour but rather in static equilibrium
%under certain conditions.
%We also rely on the vascularized model of the parenchyma proposed by
%Peterl\'{i}k et al. \cite{Peterlik2012}.

The parenchyma is modeled using corotational finite elements~\cite{Felippa}.
While relying on linear stress-strain relationship, it allows for large displacements including rotations. 
While in the full non-linear formulation, the stiffness matrix relates the forces a $\Vec{f}$ and 
displacements $\Vec{u}$ as $\Vec{f} = \Mat{K}(\Vec{u})$, the corotational model 
requires the stiffness matrix $\Mat{K}_0$ of the system be computed only once before the simulation begins. 
Then, in each step, the motion of each element $e$ is decomposed into rigid rotation $\Mat{R}^e$ and local deformation. 
The rotations are then used to update each local element stiffness matrix as $\Mat{R}\Mat{K}_0\Mat{R}^{\top}$
whereas the deformations are used to compute the linear strain in the local corotational frame.
%The stiffness matrix $\Mat{K}$ can be expressed in terms of the deformation
%$u$ and the force as $\Vec{f} = \Mat{K}(\Vec{u}) \Vec{u}$. 
%Where the
%components of $\Mat{K}$ depend on the orientation of each element in the
%current simulation step.
%the rigid component is  
% The corotational
%formulation allows us to deal with geometrical non-linearities but still
%relies on linear stress-strain relationship. Thus rotations and large
%deformations are modelled correctly. 
%In the corotational formulation the deformation
%of each element is expressed in the local frame of reference. 
There are several ways of computing the corotational frame for elements; we rely on
the geometrical method proposed in \cite{Nesme2005}.
% NOTE: This description of corotational method is very simplified and could be extended.

The model of the vascularization is based on linear beams 
with local frames of reference; in many aspects it's similar to the
corotational formulation described above~\cite{Duriez}. As such the model also handles geometric
non-linearities in the deformation. Through a specification of cross section and moments of inertia, 
the model can account for the specific properties of the blood vessels.

The beam-based model of vessels is mechanically coupled to the parenchyma as described in~\cite{peterlik2012}. 
The coupling assumes that there is no relative motion between the vessels and surrounding parenchyma. 
%The proposed model assumes no relative
%motion of vessels to the parenchyma. Segmented vascular network is
%discretized into the set of linear elements that are then projected into
%the tetrahedral mesh of the parenchyma. 
During the simulation, the nodes of linked beams are first displaced and rotated according to the actual motion of associated tetrahedra. 
As the deformation of beams results in mechanical response represented by forces and torques, these are propagated back to 
the tetrahedral FE model.

%a displacement of tetrahedral nodes are mapped  forces and torques resulting from the deformations of vessels by parenchyma are
%computed and propagated back onto the tetrahedral mesh.
